Opportunities and precautions for the use of MPAs in bycatch reduction

### 2021-12-02 documents/03_marlin_paper.Rmd
back to main

---
title: "Opportunities and precautions for the use of MPAs in bycatch reduction"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Lennon Thomas
    - Echelle Burns
date: "`r Sys.Date()`"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
  bookdown::html_document2: default
  bookdown::pdf_document2: default
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
bibliography: references.bib
params:
  run_name: ["v1.1"]
---

@@

walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv))

@@

@@

Specifically, the potential benefits of MPAs as a bycatch management are
complicated by two key issues: effort displacement and technical
interactions. One of the primary questions around MPAs is to what extent
do MPAs remove versus displace fishing effort. If MPAs remove fishing
effort, then their direct effects on target and bycatch species can only
be positive. If however MPAs displace fishing effort evaluating their
potential impacts becomes more complicated. The net conservation impact
of MPAs becomes a balancing act between the effects of reduced fishing
inside protected borders and increased fishing intensity outside MPAs.
The outcomes of this displacement can be further complicated by
technical interactions, i.e. the degree to which fishing effort for one
species affects another, particularly when species have markedly
different habitat distributions. For example, @abbott2012 demonstrated
that an MPA designed to protect red king crab (*Paralithodes
camtschaticus*) caused an unintended increase in bycatch of Pacific
halibut (*Hippoglossus stenolepis*) by the displaced bottom-trawl
fishery. Policy leakage is not a unique feature to MPAs, e.g. XX, but
the potential impacts of MPAs on target and bycatch species taking into
account the potential for these leakage effects has received relativelly
little generalized attention in the literature [see @abbott2012;
@savina2008; @fulton2011a; @baum2003 for notable exceptions].

This study addresses this gap using a novel bio-economic simulation
model that allows for efficient evaluation of the effects of MPAs on
target and bycatch species across a range of ecological and economic
dynamics, critically accounting for the potential impacts of effort
displacement and technical interactions. We use this model first to
explore general conditions that lead to positive and negative outcomes
for target and bycatch species from MPAs. We then present a case study
tuned to reflect the dynamics of the tuna, billfish, and shark
populations and fisheries of the Western and Central Pacific Ocean
(WCPO) to explore the value of optimal MPA design in managing tradeoffs
between maximizing bycatch conservation and minimizing economic impacts.
We chose the WCPO as a case study as the modern science of MPAs was
largely developed around coastal ecosystems, often characterized by
relatively static features such as coral reefs [citation]. However,
large MPAs are increasingly being implemented and considered across
pelagic ecosystems. These pelagic systems are characterized by highly
mobile species and fishing fleets using a diverse mix of static and
dynamic ocean features [@game2009; @lubchenco2003]. These ecosystems are
also home to combinations of relatively abundant species targeted by
fishing, such as many tuna stocks [@pons2016], and highly threatened
species caught as incidental "bycatch" in targeted fishing operations,
such as oceanic whitetip sharks [@pacoureau2021]. These dynamics,
combined with the increased policy interest in large MPAs covering
expanses of blue-water ecosystems, make pelagic systems such as the WCPO
an ideal case study in bycatch management through MPAs. The goal of this
paper is to provide a strategic evaluation of the sensitivity of bycatch
populations and targeted fisheries to key unknowns and design choices in
MPA planning.

@@

Along with distance, net movement is defined by habitat quality. Habitat
quality defines the quailty for adults or larvae of a given patch. The
net probability of movement from patch *i* to patch *j* is defined then
by the product of distance and habitat quality, normalized to sum to one
from each patch, such that all else being equal organisms will move the
most to the closest patch with the best habitat. We allow for seasonal
movement in the model, where years are broken into two seasons. Movement
in season two is defined as the opposite of movement in season one,
intended to reflect the behavior of seasonal migrations.

@@

Where *scale* rescales the values to be between 0 and 1 such that the
weight parameter is comparable between vectors. The algorithm works by
first assigning every patch in the system to an MPA. For the first
round, the algorithm then calculates the equilibrium objective function
value resulting from sequentially removing each individual patch from
the MPA network. The cell whose removal from the MPA produces the
highest objective function value is then removed from the MPA network,
and the marginal value of the objective function each of the remaining
cells from the last steps are then stored. For the next steps, we
sampled 25% of the remaining MPA cells with probability proportional to
their marginal contribution to the objective function in the prior
iteration, selecting the cell whose removal produces the highest total
objective function as the cell to remove from the network, and updateing
the marginal value of the sampled-but-not-selected cells. This in
essence is a sampling-importance-resampling algorithm. The algorithm
continues until no more cells are left in the MPA. This then provides
for a given objective weighting value $\Phi$ the near-optimal MPA
network at every MPA size from 0 to 100%. We say "near" as they is some
path dependency here, in that for example the network that maximizes the
objective function with 40% of the MPAs does not necessarily contain the
optimal network at 30%. However, the reversed nature of the algorithm,
starting at 100% MPA coverage and working better, produces a more stable
network with regards to this problem than starting at 0% and growing to
100%.

@@

For the generalized experiments we only include one fishing fleet. We
asign a price per KG of \$XX for tuna, intended to reflect a low value
high volume fishery, \$XX for marlin for a higher value low volume
fishery, and \$0 for sharks to reflect at true bycatch species. We
assign a selectivity of 60% of length at maturity for skipjack to allow
for some overfishing when F/M is high enough, 70% for mariln, and 90%
for shark.

We ran `r nrow(state_experiments)` state experiments, varing the state
of nature experienced by a simulation. The candidate states include

| State                 | Distribution          | Description                                     |
|-----------------------|-----------------------|-------------------------------------------------|
| $\sigma_{centroid}$   |                       | Standard deviation of habitat centroid          |
| $\sigma_{habitat}$    |                       | Standard deviation around centroid              |
| Spatial catchability  |                       | Catchability varies in sapce                    |
| Spatial allocation    |                       | Fleet allocation strategy                       |
| F/M                   |                       | Fishing mortality relative to natural mortality |
| Adult movement rate   |                       |                                                 |
| Recruit movement rate |                       |                                                 |
| Hyperallometry        |                       | Presence of hyperallometric fec                 |
| DD timing             |                       | Timing of density dependence                    |
| Seasonal Movement     | $\subset[TRUE,FALSE]$ | Seasonal Movement Present                       |

@@

We simulate degrees of species overlap using two parameters, $\sigma_c$
and $\sigma_h$. All species start with the assumption that the centroid
of their preferred habitat occurs in the center-most patch of teh two
dimensional system simulated by marlin. For our simulations, we then
generated a random deviation from that default centroid for each species
using $\sigma_c$ in the manner

$$centroid_s \sim normal(center,\sigma_c)$$ where *s* denotes species,
and *center* is the coordinate of the central cell. In this way, the
larger $\sigma_c$ grows, the more variation there is betwen the centroid
of the preferred habitat for each species.

We then control the spread of a creatures preferred habitat around that
centroid per a Gaussian dispersal kernal in the manner

@@

Where $d$ is the euclidian distance from the $centroid_s$ to patch $p$
and $\sigma_h$ controls the dispersal of the habitat around that range.

@@

One clear caveat across all of these results is that we have made two
very strong assumptions around the behavior of the fishing fleet. The
first is that we assume that effort is displaced, not removed, by the
creation of MPAs. This is a widely debated question around MPAs, and the
real impacts of MPAs on fishing effort will clearly be setting
dependent. For example, large-scale pelagic fishing fleets already have
high rates of vessel movement built into their economic models, and so
may be more likely to move rather than exit the fishery in response to
MPAs. In contrast, smaller scale coastal fishing operations may simply
exit the fishery if their local fishing grounds are placed in MPAs.
There is clear evidence of effort displacement in many fisheries though
(citation), and moreover a motivation for this study was to evaluate the
performance of MPAs in reducing bycatch given the potential for effort
displacement and technical interactions. If MPAs do not produce
increases in fishing effort outside their borders, then the direct
conservation effects of MPAs have to be positive for all species
concerned.

@@

<!-- what we found was a series of neccesary but not sufficient conditions  -->

@@

Looking at the full
`r n_distinct(total_experiment_results$placement_id) * n_distinct(total_experiment_results$state_id)`
experiments run, the results were highly variable. Across our
simulations, gains were much more common than losses, with losses of any
magnitude occurring in `r scales::percent(mean(critter_results$loss))`
of simulations with MPAs, but losses of 10% or greater of SSB in only
`r scales::percent(mean(critter_results$real_loss))` of simulations.
However we stress again that we can assign no probabilities to any of
these outcomes in the real world; this simply tells us that under the
range of conditions, selected to represent plausible states of nature,
there are many more ways for MPAs to provide a conservation effect
greater than or equal to 0 than there are for them to provide a net loss
for some species.

The question of interest then are which of the considered drivers appear
to be the most influential in affecting the magnitude and direction of
conservation effects of MPAs. Fishing mortality is the most obvious
answer; if fishing mortality is zero in the absence of an MPA then
clearly a no-take MPA will have no effect. As such we do not focus on
that metric here, focusing instead on other key drivers identified by
our simulation. Four visually prominent factors were MPA size, species,
design strategy, and degree of habitat difference. The degree to which
MPAs of a given size impact spawning biomass depends on a host of
factors. Any potential for conservation losses are intuitively convex in
MPA size: effects of any kind are low when MPA size is low, and
conservation losses must be zero when MPA size is 100% (which this graph
does not quite reach), with peak losses in our simulations occurring
between MPA sizes of XX and XX. Life history plays a clear role our
stylized with shark and marlin populations being much more likely to
experience net conservation losses as a function of MPA placement than
the more resilient tunas. High degree of habitat differences can lead to
both to the potential for large conservation gains or losses depending
on whether a more isolated species is located primarily inside or
outside of MPAs (Fig.\@ref(fig:fig-exp-biodiv)).

@@

However, this result is somewhat misleading (xx so should we just
remove?xx). Fig.\@ref(fig:fig-exp-biodiv) shows the clearly bimodal
nature of for example habitat differences. This analysis helps inform us
that habitat differences are a necessary but not sufficient condition
for conservation losses induced by effort displacement. The regression
based results require assuming that each of the effects are independent
and linear in nature. In fact each of these factors can interact in
complex and non-linear ways that may not be apparent visually or through
this type of regression analysis. To address this, we ran a random
forest using the `ranger` package (citation) in R (citation) and
extracted the impurity corrected importance scores for each of the
variables. While this does not tell us the expected marginal effect of
individual drivers, it does tell us which variables are important if our
goal is to accurately predict the potential for unintended consequences
or magnitude of conservation gains in MPAs. Note here that variables
that intuitively should have an impact but have no visually or
statistically (in a linear model) clear impact on the probability of
experiencing a conservation loss such as the dispersal distances of
adults or larvae have importance scores comparable in magnitude to
habitat differences when considered in the context of nonlinear effects
and interactions with other variables
(Fig.\@ref(fig:fig-loss-drivers)-B).

@@

MPAs can be a powerful tool for providing conservation benefits across
large swaths of species and space. Under the appropriate circumstances
they can also provide benefits to food security and economic outcomes.
However, while spatial closures are simple in concept, predicting the
likely effects of MPAs across groups of species and fishing fleets
becomes increasingly complicated as the systems being studied deviate
from assumptions of homogenously distributed species with perfectly
selective fleets. In this paper we used a new bio-economic model to
demonstrate how a range of variables often omitted in MPA modeling can
act to influence the outcomes of MPAs on conservation and economic
objectives, particularly with regards to protecting bycatch species. We
demonstrate that unintended conservation losses to both target and
bycatch species are possible, but only under a range of specific
conditions, and under most simulated cases did not occur. MPAs were
capable of causing small losses in biomass for the WCPO case study, but
only for a subset of design strategies and species, most of which could
be overcome by optimized design.

@@

The MPA literature is highly dominated by modeling studies (citation).
What then is the added value of the model and results presented here?
Our model contains many features missing from many general MPA models,
including size structure, seasonal movement, larval and adult habitats
and dispersal, and critically technical interactions xx clearer term for
this xx across species. A reader might reasonably ask about a range of
other factors not included, such as time-varying parameters and species
interactions. One could of course continue to add features to the model
until it reaches the complexity of an end-to-end model such as Atlantis
(citation). While such models have tremendous value, our goal here is to
present a model that can help provide closer to tactical advice for
users without the data or capacity for such high resolution models. One
justification for simpler MPA models is often computational
tractability: tracking the full age structure of multiple species and
the economic outcomes of multiple fleets dynamically across a high
resolution spatial grid can quickly become computationally expensive.
Faster implementations completely written in languages such as C++
however become inaccessible to many users. Our model seeks to fill a
middle ground here, with the core operations run in C++ while providing
a relatively user-friendly front end in the open-source and commonly
used R programming language. This allows users to ask the sorts of
questions addressed in this paper that require a more realistic model,
without requiring inordinate amounts of data, programming expertise, or
computational time.

Several results from our model may seem counterintuitive based on
existing literature. We found little effect of the timing and spatial
structure of density dependence in our model. Other studies have pointed
out that decisions about whether density dependence is local or global
can have major implications for fishery outcomes of MPAs (citations).
The difference here is that our model is fully age structured. As such,
yields are separated from recruitment, the location of density dependent
processes in our model, by growth, natural mortality, and adult movment.
In contrast biomass dynamics models assume that density dependence
occurs on a net "growth" parameter combining both recruitment and
biomass growth, and that this growth is instantly available to the
fishery. As such we suspect that the timing of density dependence plays
a much bigger role in the expected outcomes of MPAs when biomass
dynamics models are used as opposed to when it is included as an effect
on recruitment in an age structured model.

@marshall2021 pointed out that many fisheries models assume that
fecundity is simply proportional to spawning biomass, when in fact there
is ample empirical evidence that fecundity follows a hyperallometric
pattern, i.e. that fecundity increases faster than weight. In that study
the authors found that the presence of hyperallometric growth
substantially amplified the predicted benefits of spatiao-temporal
closures such as MPAs, relative to an assumption of allometric
fecundity. In contrast, we find that the presence of hyperallometric
fecundity, even at levels greater than the mean values reported in
@marshall2021, had little effect on the expected conservation outcomes
of MPAs (Fig.\@ref(fig:fig-loss-drivers), see SI for additional
analyses). The key distinction here is that @marshall2021 evaluated the
effect of MPAs under different assumptions of management, e.g.
management based on spawning potential ratio values calculated assuming
allmetric fecundity when in fact the relationship was hyperallometric.
Our study holds effort constant, and so reflects more the isolated
effect of hyperallometric on MPA outcomes independent of management
actions.

@rassweiler2012 and @rassweiler2014 both demonstrated the potential
importance of optimized MPA planning over more "rule of thumb" MPA
designs. In contrast, we found little value of optimization, with the
"Depletion" MPA allocation strategy more or less following the Pareto
efficiency frontier (Fig.\@ref(fig:fig-cs-totals)). As pointed out in
@rassweiler2014, the value of optimization (at least if the objective
function depends on both fishery yields and fish biomass) depends on the
degree of overfishing present. In our WCPO case study, best available
evidence is that none of the major tuna stocks that provide the bulk of
yield, revenues, and profits to the WCPO fleets are overfished or
experiencing overfishing (citations, need to update with assessments
since pons is out of date now). Our results then are consistent with
@rassweiler2014, and evaluations of our model on stocks experiencing
overfishing will likely find much higher value of optimization.

Value of information and fleet dynamics are two critical areas for
future study in this realm. Our study assumes perfect information in the
design of MPAs. We feel this is a useful benchmark for isolating the
potential impacts of key ecological and economic drivers of MPA effects.
However it is clearly not a realistic assumption, particularly for
pelagic ecosystems where species inhabit a complex, dynamic, and hard to
observe mosaic of habitats. This is likely to be particularly true for
rarer bycatch species such as sharks that may be the impetus behind many
pelagic MPA designs. Further research is needed to examine how the types
of dynamics explored here are affected by varying degrees of information
quality.

We focused our efforts on modeling the spatial distribution of a fishing
fleet conditional on a fixed pool of fishing effort. As discussed
previously, the outcomes of MPAs depend critically on the reaction of
fishing fleets to new spatial protections. In reality neither perfect
attrition nor perfect redistribution are likely to be realistic
scenarios. Instead individual fishing operations will adjust their
behavior according to their own objective functions. More realistic
models of total effort could include simple open-access dynamics in the
manner of @cabral2019, to more complex dynamics reflecting combinations
of policies such as effort controls, spatial closures, and rights based
fisheries management.

As demonstrated here there though there is a possibly endless list of
refinements and realism that could be added to simulation based studies.
While efforts such as this can help illustrate which drivers may be the
most important in particular circumstances in many cases as demonstrated
here the effect of specific variables are likely to be so context
dependent as to defy easy classification of "important" or "not". We
feel these models are better viewed as a foundation for empirical
action. In terms of planning, models such as this can be used as tool
for answering "what if" rather "what will be". Helping community members
visualize the potential impacts of alternative states of nature, such as
debates around life history or habitat distribution, can help set
expectations and guide data collection. Models such as this can also
serve as hypotheses for expected effects that can ideally be confronted
with data as MPAs are actually implemented.

@@

>>>>

---
title: "Opportunities and precautions for the use of MPAs in bycatch reduction"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Lennon Thomas
    - Echelle Burns
date: "`r Sys.Date()`"
output: 
  bookdown::word_document2:
    number_sections: false
    reference_docx: template.docx
  bookdown::html_document2: default
  bookdown::pdf_document2:
    number_sections: false
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
bibliography: references.bib
params:
  run_name: ["v1.1"]
csl: proceedings-of-the-royal-society-b
---

@@

walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv)) # load results into global environment

@@

# Abstract

Marine protected areas (MPAs) are a growing part of the marine
conservation toolbox. One of the goals of global MPA expansion is to protect and restore biodiversity through conservation of threatened species often caught incidentally in fisheries, such as marine mammals, turtles, seabirds, and sharks. However, the ability of MPAs to provide population-level conservation benefits to bycatch species depends in part on the ways in which fishing fleets react to MPA placement, and the degree to which habitat and movement dynamics of different species within an ecosystem overlap. In this paper we assess the impacts of MPAs on target and bycatch species and fisheries when faced with fleet redistribution and heterogeneity in life history and habitat use using a novel bio-economic model. We show that while in the majority (xx%) of simulated cases MPAs provided net conservation benefits to all species, in xx% of simulations MPAs resulted in net losses in biomass for 1 or more species. While factors interact in complex and nonlinear ways, placement strategy, adult and larval movement, and degree of habitat heterogeneity were among the most important factors in determining whether MPAs produced a net conservation loss. We then used our model to assess the potential impacts of MPAs in the Western and Central Pacific Ocean on tuna, billfish, and shark populations and fisheries. MPAs produced net conservation gains for all species in nearly all simulations, but did so at the expense of both yields and profits, even when MPA design was optimized. Our results demonstrate the importance of considering fleet behavior, movement dynamics, and habitat heterogeneity in assessing the potential impacts of MPAs in a multi-species and multi-fleet context. XX some punchier closing sentence XX.

# To Do

- Produce results that design MPAs around only a subset of species, to see how sensitive results are to say not including some species in the analysis

- Expand more on the WCPO case study and make the paper a bit more *about* that: make clearer how the model could be used in a real-world setting

- Link better to Will White and company's work on MPA modeling etc

@@

The potential benefits of MPAs as a bycatch management are
complicated by two key issues: effort displacement and technical
interactions. One of the primary questions around MPAs is to what extent
do MPAs remove versus displace fishing effort. If MPAs remove fishing
effort, then their direct effects on target and bycatch species can only
be positive. If however MPAs displace fishing effort, evaluating their
potential impacts becomes more complicated. The net conservation impact
of MPAs becomes a balancing act between the effects of reduced fishing
inside protected borders and increased fishing intensity outside MPAs.
The outcomes of this displacement can be further complicated by
technical interactions, i.e. the degree to which fishing effort for one
species affects another, particularly when species have markedly
different habitat distributions. For example, @abbott2012 demonstrated
that an MPA designed to protect red king crab (*Paralithodes
camtschaticus*) caused an unintended increase in bycatch of Pacific
halibut (*Hippoglossus stenolepis*) by the displaced bottom-trawl
fishery. Policy leakage is not a unique feature to MPAs, e.g. XX include non-marine examples XX, but
the potential impacts of MPAs on target and bycatch species taking into
account the potential for these leakage effects has received relatively
little generalized attention in the literature [see @abbott2012;
@savina2008; @fulton2011a; @baum2003 for notable exceptions].

We address this gap by providing a strategic evaluation of the sensitivity of bycatch
populations and targeted fisheries to key unknowns and design choices in
MPA planning. We used a novel bio-economic simulation
model that allows for efficient evaluation of the effects of MPAs on
target and bycatch species across a range of ecological and economic
dynamics, critically accounting for the potential impacts of effort
displacement and technical interactions.  We used this model first to
explore general conditions that lead to positive and negative outcomes
for target and bycatch species from MPAs. We then present a case study
tuned to reflect the dynamics of the tuna, billfish, and shark
populations and fisheries of the Western and Central Pacific Ocean
(WCPO) to explore the value of optimal MPA design in managing tradeoffs
between maximizing bycatch conservation and minimizing economic impacts.

We chose the WCPO as a case study as the modern science of MPAs was
largely developed around coastal ecosystems, often characterized by
relatively static features such as coral reefs [citation]. However,
large MPAs are increasingly being implemented and considered across
pelagic ecosystems [@hilborn2021c]. These pelagic systems are characterized by highly
mobile species and fishing fleets using a diverse mix of static and
dynamic ocean features [@game2009; @lubchenco2003]. These ecosystems are
also home to combinations of relatively abundant species targeted by
fishing, such as many tuna stocks [@pons2016], and highly threatened
species caught as incidental "bycatch" in targeted fishing operations,
such as oceanic whitetip sharks [@pacoureau2021]. These dynamics,
combined with the increased policy interest in large MPAs covering
expanses of blue-water ecosystems, make pelagic systems such as the WCPO
an ideal case study in bycatch management through MPAs.

@@

Along with distance, net movement is defined by habitat quality. Habitat
quality defines the quality for adults or larvae of a given patch. The
net probability of movement from patch *i* to patch *j* is defined then
by the product of distance and habitat quality, normalized to sum to one
from each patch, such that all else being equal organisms will move the
most to the closest patch with the best habitat. We allow for seasonal
movement in the model, where years are broken into two seasons. Movement
in season two is defined as the opposite of movement in season one,
intended to reflect the behavior of seasonal migrations.

@@

Where *scale* rescales the values to be between 0 and 1 such that the
weight parameter is comparable between vectors. The algorithm works by
first assigning every patch in the system to an MPA. For the first
round, the algorithm then calculates the equilibrium objective function
value resulting from sequentially removing each individual patch from
the MPA network. The cell whose removal from the MPA produces the
highest objective function value is then removed from the MPA network,
and the marginal value of the objective function each of the remaining
cells from the last steps are then stored. For the next steps, we
sampled 25% of the remaining MPA cells with probability proportional to
their marginal contribution to the objective function in the prior
iteration, selecting the cell whose removal produces the highest total
objective function as the cell to remove from the network, and updating
the marginal value of the sampled-but-not-selected cells. This in
essence is a sampling-importance-resampling algorithm. The algorithm
continues until no more cells are left in the MPA. This then provides
for a given objective weighting value $\Phi$ the near-optimal MPA
network at every MPA size from 0 to 100%. We say "near" as they is some
path dependency here, in that for example the network that maximizes the
objective function with 40% of the MPAs does not necessarily contain the
optimal network at 30%. However, the reversed nature of the algorithm,
starting at 100% MPA coverage and working better, produces a more stable
network with regards to this problem than starting at 0% and growing to
100%.

@@

For the generalized experiments we only include one fishing fleet. We
assign a price per KG of \$XX for tuna, intended to reflect a low value
high volume fishery, \$XX for marlin for a higher value low volume
fishery, and \$0 for sharks to reflect at true bycatch species. We
assign a selectivity of 60% of length at maturity for skipjack to allow
for some overfishing when F/M is high enough, 70% for marlin, and 90%
for shark.

We ran `r nrow(state_experiments)` state experiments, varying the state
of nature experienced by a simulation. The candidate states include

| State                 | Distribution          | Description                                     |
|-----------------------|-----------------------|-------------------------------------------------|
| $\sigma_{centroid}$   |                       | Standard deviation of habitat centroid          |
| $\sigma_{habitat}$    |                       | Standard deviation around centroid              |
| Spatial catchability  |                       | Catchability varies in space                    |
| Spatial allocation    |                       | Fleet allocation strategy                       |
| F/M                   |                       | Fishing mortality relative to natural mortality |
| Adult movement rate   |                       |                                                 |
| Recruit movement rate |                       |                                                 |
| Hyperallometry        |                       | Presence of hyperallometric fec                 |
| DD timing             |                       | Timing of density dependence                    |
| Seasonal Movement     | $\subset[TRUE,FALSE]$ | Seasonal Movement Present                       |

@@

We simulate degrees of species overlap using two parameters, $\sigma_c$
and $\sigma_h$. All species start with the assumption that the centroid
of their preferred habitat occurs in the center-most patch of the two
dimensional system simulated by marlin. For our simulations, we then
generated a random deviation from that default centroid for each species
using $\sigma_c$ in the manner

$$centroid_s \sim normal(center,\sigma_c)$$ where *s* denotes species,
and *center* is the coordinate of the central cell. In this way, the
larger $\sigma_c$ grows, the more variation there is between the centroid
of the preferred habitat for each species.

We then control the spread of a creatures preferred habitat around that
centroid per a Gaussian dispersal kernel in the manner

@@

Where $d$ is the euclidean distance from the $centroid_s$ to patch $p$
and $\sigma_h$ controls the dispersal of the habitat around that range.

@@

One clear caveat across all of these results is that we have made two
very strong assumptions around the behavior of the fishing fleet. The
first is that we assume that effort is displaced, not removed, by the
creation of MPAs. This is a widely debated question around MPAs [@hilborn2021c], and the
real impacts of MPAs on fishing effort will clearly be setting
dependent. For example, large-scale pelagic fishing fleets already have
high rates of vessel movement built into their economic models, and so
may be more likely to move rather than exit the fishery in response to
MPAs. In contrast, smaller scale coastal fishing operations may simply
exit the fishery if their local fishing grounds are placed in MPAs.
There is clear evidence of effort displacement in many fisheries though
(citation), and moreover a motivation for this study was to evaluate the
performance of MPAs in reducing bycatch given the potential for effort
displacement and technical interactions. If MPAs do not produce
increases in fishing effort outside their borders, then the direct
conservation effects of MPAs have to be positive for all species
concerned.

@@

<!-- what we found was a series of necessary but not sufficient conditions  -->

@@

Looking at the full
`r n_distinct(total_experiment_results$placement_id) * n_distinct(total_experiment_results$state_id)`
experiments run gains were much more common than losses, with losses of any
magnitude occurring in `r scales::percent(mean(critter_results$loss))`
of simulations with MPAs, but losses of 10% or greater of SSB in only
`r scales::percent(mean(critter_results$real_loss))` of simulations.
However we stress again that we can assign no probabilities to any of
these outcomes in the real world; this simply tells us that under the
range of conditions, selected to represent plausible states of nature,
there are many more ways for MPAs to provide a conservation effect
greater than or equal to 0 than there are for them to provide a net loss
for some species.

The question of interest then are which of the considered drivers appear
to be the most influential in affecting the magnitude and direction of
conservation effects of MPAs. Fishing mortality is the most obvious
answer; if fishing mortality is zero in the absence of an MPA then
clearly a no-take MPA will have no effect. As such we do not focus on
that metric here, focusing instead on other key drivers identified by
our simulation. Four visually prominent factors were MPA size, species,
design strategy, and degree of habitat difference. The degree to which
MPAs of a given size impact spawning biomass depends on a host of
factors. Any potential for conservation losses are intuitively convex in
MPA size: effects of any kind are low when MPA size is low, and
conservation losses must be zero when MPA size is 100% (which this graph
does not quite reach). Peak losses in our simulations occurring
between MPA sizes of XX and XX. Life history plays a clear role our
stylized with shark and marlin populations being much more likely to
experience net conservation losses as a function of MPA placement than
the more resilient tunas. High degree of habitat differences can lead to
both to the potential for large conservation gains or losses depending
on whether a more isolated species is located primarily inside or
outside of MPAs (Fig.\@ref(fig:fig-exp-biodiv)).

@@

However, this result is somewhat misleading (xx so should we just
remove?xx). Fig.\@ref(fig:fig-exp-biodiv) shows the clearly bimodal
nature of for example habitat differences. This analysis helps inform us
that habitat differences are a necessary but not sufficient condition
for conservation losses induced by effort displacement. The regression
based results require assuming that each of the effects are independent
and linear in nature. In fact each of these factors can interact in
complex and non-linear ways that may not be apparent visually or through
this type of regression analysis. To address this, we ran a random
forest using the `ranger` package [@wright2017] in R [@rcoreteam2021] and
extracted the impurity corrected importance scores for each of the
variables. While this does not tell us the expected marginal effect of
individual drivers, it does tell us which variables are important if our
goal is to accurately predict the potential for unintended consequences
or magnitude of conservation gains in MPAs. Note here that variables
that intuitively should have an impact but have no visually or
statistically (in a linear model) clear impact on the probability of
experiencing a conservation loss such as the dispersal distances of
adults or larvae have importance scores comparable in magnitude to
habitat differences when considered in the context of nonlinear effects
and interactions with other variables
(Fig.\@ref(fig:fig-loss-drivers)-B).

@@

MPAs can be a powerful tool for providing conservation benefits across
large swaths of species and space. Under the appropriate circumstances
they can also provide benefits to food security and economic outcomes.
However, while spatial closures are simple in concept, predicting the
likely effects of MPAs across groups of species and fishing fleets
becomes increasingly complicated as the systems being studied deviate
from assumptions of homogeneously distributed species with perfectly
selective fleets. In this paper we used a new bio-economic model to
demonstrate how a range of variables often omitted in MPA modeling can
act to influence the outcomes of MPAs on conservation and economic
objectives, particularly with regards to protecting bycatch species. We
demonstrate that unintended conservation losses to both target and
bycatch species are possible, but only under a range of specific
conditions, and under most simulated cases did not occur. MPAs were
capable of causing small losses in biomass for the WCPO case study, but
only for a subset of design strategies and species, most of which could
be overcome by optimized design.

@@

The MPA literature is highly dominated by modeling studies [@oleary2016].
What then is the added value of the model and results presented here?
Our model contains many features missing from many general MPA models,
including size structure, seasonal movement, larval and adult habitats
and dispersal, and critically technical interactions xx clearer term for
this xx across species. A reader might reasonably ask about a range of
other factors not included, such as time-varying parameters and species
interactions. One could of course continue to add features to the model
until it reaches the complexity of an end-to-end model such as Atlantis [@fulton2011a]. While such models have tremendous value, our goal here is to
present a model that can help provide closer to tactical advice for
users without the data or capacity for such high resolution models. One
justification for simpler MPA models is often computational
tractability: tracking the full age structure of multiple species and
the economic outcomes of multiple fleets dynamically across a high
resolution spatial grid can quickly become computationally expensive.
Faster implementations completely written in languages such as C++
however become inaccessible to many users. Our model seeks to fill a
middle ground here, with the core operations run in C++ while providing
a relatively user-friendly front end in the open-source and commonly
used R programming language. This allows users to ask the sorts of
questions addressed in this paper that require a more realistic model,
without requiring inordinate amounts of data, programming expertise, or
computational time.

Several results from our model may seem counterintuitive based on
existing literature. We found little effect of the timing and spatial
structure of density dependence in our model. Other studies have pointed
out that decisions about whether density dependence is local or global
can have major implications for fishery outcomes of MPAs (citations).
The difference here is that our model is fully age structured. As such,
yields are separated from recruitment, the location of density dependent
processes in our model, by growth, natural mortality, fishery selectivity, and adult movement.
In contrast biomass dynamics models assume that density dependence
occurs on a net "growth" parameter combining both recruitment and
biomass growth, and that this growth is instantly available to the
fishery. As such we suspect that the timing of density dependence plays
a much bigger role in the expected outcomes of MPAs when biomass
dynamics models are used as opposed to when it is included as an effect
on recruitment in an age structured model.

@marshall2021 pointed out that many fisheries models assume that
fecundity is simply proportional to spawning biomass, when in fact there
is ample empirical evidence that fecundity follows a hyperallometric
pattern, i.e. that fecundity increases faster than weight. In that study
the authors found that the presence of hyperallometric growth
substantially amplified the predicted benefits of spatio-temporal
closures such as MPAs, relative to an assumption of allometric
fecundity. In contrast, we find that the presence of hyperallometric
fecundity, even at levels greater than the mean values reported in
@marshall2021, had little effect on the expected conservation outcomes
of MPAs (Fig.\@ref(fig:fig-loss-drivers), see SI for additional
analyses). The key distinction here is that @marshall2021 evaluated the
effect of MPAs under different assumptions of management, e.g.
management based on spawning potential ratio values calculated assuming
allometric fecundity when in fact the relationship was hyperallometric.
Our study holds effort constant, and so reflects more the isolated
effect of hyperallometric on MPA outcomes independent of management
actions.

@rassweiler2012 and @rassweiler2014 both demonstrated the potential
importance of optimized MPA planning over more "rule of thumb" MPA
designs. In contrast, we found little value of optimization, with the
"Depletion" MPA allocation strategy more or less following the Pareto
efficiency frontier (Fig.\@ref(fig:fig-cs-totals)). As pointed out in
@rassweiler2014, the value of optimization (at least if the objective
function depends on both fishery yields and fish biomass) depends on the
degree of overfishing present. In our WCPO case study, best available
evidence is that none of the major tuna stocks that provide the bulk of
yield, revenues, and profits to the WCPO fleets are overfished or
experiencing overfishing (citations, need to update with assessments
since pons is out of date now). Our results then are consistent with
@rassweiler2014, and evaluations of our model on stocks experiencing
overfishing will likely find much higher value of optimization. As a demonstration of this, we ran a sensitivity analyses dramatically increasing the fishing mortality rates of all the tuna species in our WCPO case study. We found that indeed the value of optimization on profits increased by XX% (calculate area of the ribbon between the two curves as a measure of value of optimization).

Value of information and fleet dynamics are two critical areas for
future study. Our study assumes perfect information in the
design of MPAs. This is a useful benchmark for isolating the
potential impacts of key ecological and economic drivers of MPA effects.
However it is clearly not a realistic assumption, particularly for
pelagic ecosystems where species inhabit a complex, dynamic, and hard to
observe mosaic of habitats. This is likely to be particularly true for
rarer bycatch species such as sharks that may be the impetus behind many
pelagic MPA designs. Further research is needed to examine how the types
of dynamics explored here are affected by varying degrees of information
quality.

We focused our efforts on modeling the spatial distribution of a fishing
fleet conditional on a fixed pool of fishing effort. The outcomes of MPAs depend critically on the reaction of
fishing fleets to new spatial protections. In reality neither perfect
attrition nor perfect redistribution are likely to be realistic
scenarios. Instead individual fishing operations will adjust their
behavior according to their own objective functions. More realistic
models of total effort could include simple open-access dynamics in the
manner of @cabral2019, to more complex dynamics reflecting combinations
of policies such as effort controls, spatial closures, and rights based
fisheries management.

There is a possibly endless list of
refinements and realism that could be added to simulation based studies.
While efforts such as this can help illustrate which drivers may be the
most important in particular circumstances in many cases as demonstrated
here the effect of specific variables are likely to be so context
dependent as to defy easy classification of "important" or "not". These models are better viewed as a foundation for empirical
action. In terms of planning, models such as this can be used as tool
for answering "what if" rather "what will be". Helping community members
visualize the potential impacts of alternative states of nature, such as
debates around life history or habitat distribution, can help set
expectations and guide data collection. Models such as this can also
serve as hypotheses for expected effects that can ideally be confronted
with data as MPAs are actually implemented.

@@

\renewcommand{\thesection}{Appendix S\arabic{section}}

\renewcommand{\thefigure}{S\arabic{figure}}
